---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bpcandata

The goal of bpcandata is to make StatCan public use microdata files easier to access and work with in R. 

## Installation

You can install the development version of bpcandata from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("context-dependent/bpcandata")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(bpcandata)
library(tidyverse)
```

## Loading the data

`fetch_lfs_pumf` downloads the LFS PUMF files for a given year (or vector of multiple years) and extracts them to a specfied `cache_dir`. 
In this case, we download the 2015 LFS PUMF files to the `"./local"` directory.
`read_lfs_pumf` reads the LFS PUMF files from the specified directory, returning a list of two data frames: `records` and `codebook`.

```{r load_data}
fetch_lfs_pumf(2015, cache_dir = "local")
d15 <- read_lfs_pumf("local/2015")

d15$records |> head()

d15$codebook |> head()
```

To view the codebook in a human readable format, run `?bpcandata::read_lfs_pumf` in the R console. 

## Applying the codebook

```{r apply_codebook}
d15_labelled <- encode_lfs_factors(d15$records, d15$codebook)

d15_labelled |> head()
```

`encode_lfs_factors` applies the codebook to the records data frame, converting the factor codes to their corresponding labels.

## Point estimates using FINALWT

If you aren't concerned with sampling variability, you can use the `FINALWT` column as frequency weights to calculate totals, means, and proportions.

```{r point_estimates}
d15_labelled |>
  group_by(SURVMNTH, LFSSTAT) |>
  summarize(
    n = sum(FINALWT)
  ) |>
  pivot_wider(names_from = LFSSTAT, values_from = n)
```

## Estimating sampling variability with bootstrap weights

If sampling variability is important, generate appropriately calibrated bootstrap replicate weights with `generate_lfs_bootstrap_weights`.

```{r generate_weights}
repwt15 <- generate_lfs_bootstrap_weights(d15_labelled, 100)
```

## Tabulating bootstrap estimates

`bpcandata` provides functions for tabulating grouped totals and means with bootstrap replicate weights. 

```{r tabulate_bootstrap}
d15_labelled$REPWT <- repwt15

d15_labelled |>
  group_by(SURVMNTH, LFSSTAT) |>
  summarize(
    n = lfs_bs_total(REPWT, FINALWT)
  ) |>
  unpack(n, names_sep = "_") |>
  head()

d15_labelled |>
  group_by(SURVMNTH) |>
  filter(
    LFSSTAT %in% c("Unemployed", "Not in labour force"),
    EVERWORK != "No, never worked"
  ) |>
  summarize(mean_wks_away = lfs_bs_mean(DURJLESS, REPWT, FINALWT)) |>
  unpack(mean_wks_away, names_sep = "_") |>
  head()
```